Supplement 1: Different Response Distributions

### 2018-10-21 pkg/disttree/inst/draft/supplement/supplement_distributions.Rnw
recalculation of Rain_distributions.R after bugfix and adaption of the corresponding supplement
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2345 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: e6432f985fe0140900ab2a8615af671e9020aa26

\item Similarly, for the EMOS model all distributions lead to comparable CRPS for all
stations. However, there is some more variation in CRPS across distributions compared to
the distributional forest.

\item In contrast, the two GAMLSS models typically perform better (i.e., have lower CRPS) for
the Gaussian distributions compared to the logistic distribution. The preference of
censored vs.\ hurdle specification is not unanimous across stations.

@@

load("Rain_distributions.rda")
levels(results$means$distribution) <- c("cgaussian", "hgaussian", "clogistic")
levels(results$means$station)[16] <- "St. Johann im Walde"

@@

\item For EMOS, the median skill score for both Gaussian hurdle and zero-censored logistic
  are very close to the zero-censored Gaussian reference. However, the variation for the
  hurdle model is much higher than for the censored model.

\item For the two GAMLSS models the Gaussian hurdle is quite close to the zero-censored
  Gaussian reference, while the zero-censored logistic clearly performs worse (especially
  in the prespecified GAMLSS).

@@

# boxplots with matplot
par(mfrow = c(1,4), mar = c(4,4,3,0.7))
matplot(t(s_df[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "Distributional forest",
        xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_df, add = TRUE, col = "transparent")
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Distributional forest"), main = "Distributional forest",
#        ylim = c(-0.12,0.07), las = 2, ylab = "CRPS skill score")
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,2.9,3,1.8))
matplot(t(s_g[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Prespecified GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_g, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Prespecified GAMLSS"), main = "Prespecified GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,1.8,3,2.9))
matplot(t(s_gb[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Boosted GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_gb, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Boosted GAMLSS"), main = "Boosted GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,0.7,3,4))
matplot(t(s_emos[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "EMOS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_emos, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "EMOS"), main = "EMOS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_gaussian} CRPS skill score (higher = better)
for hurdle Gaussian and censored logistic distributions in reference to the
censored Gaussian distribution (orange horizontal line). Gray lines correspond to the 15~stations and are
aggregated in the boxplot. The y-axis range is limited
to $-10$\%, lower values for the two GAMLSS models are not displayed.}
\end{figure}

@@

Figures~\ref{fig:box_crpss_EMOS} and \ref{fig:box_crps} compare the methods separately 
for each distribution, once based on CRPS skill score values with reference model EMOS 
and once based on CRPS values. Both confirm the conclusion that the censored Gaussian 
model and the hurdle model lead to almost the same results for all methods while only 
the logistic model shows small variations.

Altogether it can be stated that for this application the left-censored at zero Gaussian distribution seems 
to be an appropriate choice while the other two distributions, especially the hurdle model, 
are also reasonable alternatives. However, they do not lead to an overall improvement in 
performance for this application.

@@

This supplement shows that the zero-censored Gaussian models from the main
manuscript and the alternative distributional specifications in this
supplement perform at least similarly. For some combinations of models and
stations, the Gaussian hurdle leads to small improvements on average but
typically with more variability (especially for EMOS and the boosted
GAMLSS). Therefore, the main manuscript focuses on the zero-censored
Gaussian model which is well-established in the probabilistic
precipitation forecasting literature. Moreover, the distributional forest
can be shown to be rather robust against distributional
(mis-)specifications compared to the other models.

>>>>

\item Also for the prespecified GAMLSS all three distributions lead to almost the same 
results and no clear advantage of one or the other distribution can be detected.

\item Similarly, for the EMOS model and the boosted GAMLSS all distributions 
lead to comparable CRPS for all stations. However, there is some more variation 
in CRPS across distributions compared to the distributional forest.

@@

load("Rain_distributions.rda")
levels(results$means$distribution) <- c("cgaussian", "hgaussian", "clogistic")
levels(results$means$station)[15] <- "St. Johann im Walde"

@@

\item Again, for the prespecified GAMLSS the results are very similar as for the 
  distributional forest showing only a slight advantage of the Gaussian hurdle model, 
  however, not for all stations. 
  %The zero-censored logistic distribution performs on average equally
  %well as the the zero-censored Gaussian distribution.

\item For the boosted GAMLSS, the Gaussian hurdle is quite close to the zero-censored
  Gaussian reference but shows a high variation, while the zero-censored logistic 
  performs slightly better.

\item For EMOS, the median skill score for both Gaussian hurdle and zero-censored logistic
  are very close to the zero-censored Gaussian reference. However, as for the boosted GAMLSS
  the variation for the hurdle model is much higher than for the censored model.

@@

# boxplots with matplot
par(mfrow = c(1,4), mar = c(4,4,3,0.7))
matplot(t(s_df[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "Distributional forest",
        xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 2.5),
        ylim = c(-0.05, 0.05))
boxplot(s_df, add = TRUE, col = "transparent")
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Distributional forest"), main = "Distributional forest",
#        ylim = c(-0.12,0.07), las = 2, ylab = "CRPS skill score")
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,2.9,3,1.8))
matplot(t(s_g[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Prespecified GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.05, 0.05))
boxplot(s_g, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Prespecified GAMLSS"), main = "Prespecified GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,1.8,3,2.9))
matplot(t(s_gb[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Boosted GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.05, 0.05))
boxplot(s_gb, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Boosted GAMLSS"), main = "Boosted GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,0.7,3,4))
matplot(t(s_emos[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "EMOS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.05, 0.05))
boxplot(s_emos, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "EMOS"), main = "EMOS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_gaussian} CRPS skill score (higher = better)
for hurdle Gaussian and censored logistic distributions in reference to the
censored Gaussian distribution (orange horizontal line). Gray lines correspond to the 15~stations and are aggregated in the boxplot.}
\end{figure}

@@

Figures~\ref{fig:box_crpss_EMOS} and \ref{fig:box_crps} compare the methods separately 
for each distribution, once based on CRPS skill score values with reference model EMOS 
and once based on CRPS values. Both confirm the conclusion that all three distributions
lead to almost the same results for all methods.

Altogether it can be stated that for this application the left-censored at zero Gaussian
distribution seems to be an appropriate choice while the other two distributions are also
reasonable alternatives. However, they do not lead to an overall improvement in 
performance for this application.

@@

This supplement shows that the zero-censored Gaussian models from the main
manuscript and the alternative distributional specifications in this
supplement perform at least similarly. For some combinations of models and
stations, the Gaussian hurdle leads to small improvements on average but
typically with more variability (especially for EMOS and the boosted
GAMLSS). Furthermore, for the boosted GAMLSS the zero-censored logistic
distribution shows slightly better results, however, only within a range of
less than $2\%$ and not for all stations.
Therefore, the main manuscript focuses on the zero-censored
Gaussian model which is well-established in the probabilistic
precipitation forecasting literature. Moreover, the distributional forest
can be shown to be rather robust against distributional
(mis-)specifications compared to the other models.
